---
title: "Did Fukushima nuclear accident change people's concerns about the environment?"
author: "Uni Lee"
date: "3/26/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven)
library(dplyr)
library(janitor)
library(ggplot2)
library(table1)
library(stringr)
library(stargazer)
library(plm)

```

# Introduction

In March 2011, Fukushima-daichi nuclear power plant in Japan exploded to cause a catastrophic damage to the local environment and health of the residents. It was the second largest nuclear explosion in history, following the Chernobyl accident in 1986. This catastrophic event triggered triggered a major policy change in Germany. Three months after the accident, German Chancellor Angela Merkel announced that the oldest nuclear power plants in Germany will be shut down permanently. The German parliament passed a bill that would completely shut-down all German reactors by 2022. This swift change in energy policy was unprecedented and unique in Germany history. Up until Fukushima, parties were divided on the nuclear power issue. However, news of this distant threat made the conservative government took a swift turn in their political stance and align with the liberal parties. How much impact did the Fukushima disaster have on the change in people's concerns about the environment? This research paper analyzes German Socio-Economic Panel (SOEP) data to distill the effect of Fukushima disaster on public views on environmental problems. 

# Theory and hypothesis

Judging from the increased political support for nuclear phase-out policy after the disaster, it can be hypothesized that Fukushima disaster increased people's concerns about the environment. This hypothesis is supported by Kingdon (1984)'s notion of "three streams of policy change". According to Davidsson (2020), major global disasters like Fukushima opens a window of opportunity for policy change by increasing the sense of urgency and political legitimacy for a policy action. After a major accident, network of actors related to policy-making is strengthened and cross-party collaborations allow them to reach a consensus faster (Davidsson, 2020).

Empirical studies have supported the hypothesis. Goelbel et. al (2013) have found that the Fukushima disaster significantly increased environmental concerns by 20% in Germany. The finding is consistent when applied to a similar nuclear disaster. Berger (2008) has found that the Chernobyl accident in 1986 increased the probability of person being very worried about the environment increased on average by about 16 percentage points.  

# Data and Methodology

To measure the effect of Fukushima disaster on people's worries about the environment, we will employ Fixed Effects (FE) and difference-in-difference (DiD) approach on German Socio-Economic Panel data (SOEP). For the purpose of this study, we will focus on individuals who took part in both 2010-2011. The analytical sample contains 14,394 person-years and 7,197 persons. 

FE and DiD models allow us to control for unobserved heterogeneity by taking advantage of the time dimension of the data. The fixed effects model measure the within-individual change in environmental concerns after the accident. Individuals who completed the study before the accident are dropped out of the fixed effects model. 

In the difference-in-difference model, all individuals are included in the study. The people who completed the survey before Fukushima are considered the "control" group, since they have not been exposed to the event yet. People who completed the survey after are the "treatment" group. The change in environmental concerns will be compared between the control and treatment group to measure the effect Fukushima, under the assumption of parallel time trends. This approach has been adopted from Goebel et al (2013). 

Since neither of these models can estimate the coefficients for time-constant variables, results of the data are also compared to Ordinary Least Square regression model to understand the demographic differences. 


```{r}
soep_raw <- readRDS("data/08_soep/SOEP_4.rds") %>% 
  clean_names()

# Filter out valid cases and change factors into numeric values
soep_unbal <- soep_raw %>%
  select(-plh0160) %>%
  filter(plh0036 %in% c(1,2,3),
         edu %in% c("2-medium", "1-low", "3-high"),
         emp %in% c("1-Employment", "2-Unemployment", "3-Education", "Not working"),
         satis %in% c(0,1,2,3,4,5,6,7,8,9,10),
         health %in% c(0,1,2,3,4,5,6,7,8,9,10),
         year >= 2008, year <= 2016,
         pmonin > 0) %>%
  mutate(plh0036 = as.integer(plh0036),
         mig = as.factor(mig),
         sex = as.factor(sex),
         emp = as.character(emp),
         fam = as.factor(fam),
         satis = as.numeric(satis),
         health = as.numeric(health),
         month = as.numeric(pmonin)) %>%
 mutate(plh0036 = as.factor(case_when(
          plh0036 == 1 ~ 3, # Make 3 as "most worried"
          plh0036 == 2 ~ 2,
          plh0036 == 3 ~ 1)), # Make 1 as "least worried"
        emp = as.factor(case_when(
          emp == "1-Employment" ~ emp,
          emp == "2-Unemployment" ~ emp,
          emp %in% c("3-Education", "Not working") ~ "3-Non-working")),
        agec = as.factor(case_when(
          agec %in% c("17-19", "20-29") ~ "17-29",
          agec == "30-39" ~ "30-39",
          agec == "40-49" ~ "40-49",
          agec == "50-59" ~ "50-59",
          agec %in% c("60-69", "70+") ~ "60+" 
  ))) %>%
  mutate(fukushima = as.factor(case_when(
     year <= 2010 ~ 0,
     year == 2011 & month < 3 ~ 0,
     year == 2011 & month >= 3 ~ 1,
     year >= 2012 ~ 1))) %>%
   na.omit() 

nrow(soep_unbal)
length(unique(soep_unbal$id))

# Selecting only 2010-2011
soep_unbal_1011 <- soep_unbal %>%
  filter(year %in% c(2010, 2011)) 

nrow(soep_unbal_1011)
length(unique(soep_unbal_1011$id))

# Balance the dataset - get subjects who have done the survey both in 2010 and 2011
soep_bal_1011 <- soep_unbal_1011 %>%
  group_by(id) %>%
  mutate(spell=row_number(),
         max_spell=max(spell)) %>%
  filter(max_spell==2) %>%
  ungroup() %>%
  select(-spell, -max_spell) %>%
  mutate(plh0036 = as.numeric(plh0036))

nrow(soep_bal_1011)          
length(unique(soep_bal_1011$id))

```


## Variables of Interest 

The dependent variable in this study is the level of people's worries about the environment _(plh0036)_. The SOEP questionnaire includes the following question: *"how worried are you about the environment?"* Respondents rate this question on the scale of three: 
-- 1: very worried
-- 2: somewhat worried
-- 3: not worried

The main independent variable in this study is a dummy variable named _fukushima_ that indicates whether the interview was taken place before or after Fukushima accident.
-- 0: Before March 2011
-- 1: After March 2011 (includes March)

The following demographic variables that are taken into consideration for descriptive statistics and OLS regression:

* sex: 1-Female 2-Male 
* fam: 1-not married, 2-married
* emp: 1-Employment, 2-Unemployment, 3-Education or not working       
* age group (agec), 
* education (Low, medium, high)

In addition, we account for the following time-varying individual heterogeneity in the FE and DiD models: 

* life satisfaction (0 - 10)
* health (0 - 10)


# Descriptive Statistics

## How people's concerns about the environment changes over time

```{r}
# For descriptive stats 
soep_time <- soep_raw %>%
  filter(plh0036 %in% c(1,2,3)) %>%
  group_by(year) %>%
  summarize(mean_concern = mean(plh0036))


ggplot(soep_time, aes(x=year, y=mean_concern)) +
  geom_line() + geom_vline(xintercept=2011, linetype="dashed", color="red") +
  theme_classic() +
  labs(title="Average concerns about the environment 1984-2018",
       y="average concern about the environment") +
  theme(axis.title.x=element_blank(),
        title=element_text(vjust=0.5))
```

## Demographics

```{r include = FALSE}
# Labels for the summary table
table1::label(soep_bal_1011$plh0036) <- "Environmental Concern"
table1::label(soep_bal_1011$fukushima) <- "Fukushima"
table1::label(soep_bal_1011$sex) <- "Gender"
table1::label(soep_bal_1011$mig) <- "Migration status"
table1::label(soep_bal_1011$emp) <- "Employment status"
table1::label(soep_bal_1011$fam) <- "Family statis"
table1::label(soep_bal_1011$edu) <- "Education level"

soep_2010 <- soep_bal_1011 %>%
  filter(year==2010)

soep_2011 <- soep_bal_1011 %>%
  filter(year==2011)

table1::table1(~plh0036 + fukushima + sex + mig + emp + fam + edu, data=soep_2010) 
table1::table1(~plh0036 + fukushima + sex + mig + emp + fam + edu, data=soep_2011) 


```

# Ordinary Least Squares

For a simple OLS, we regress the score for environmental concerns on Fukushima dummy variable and control for the demographic variables. The regression results show that Fukushima increased worries about the environment by 0.040-0.042 (1.4%). This result is statistically significant, but not substantial. The models also show that men are generally less worried than women. In terms of the age group, people in their 40s and 50s are most worried in comparison to other age groups. Higher education leads to higher level of environmental concerns. Since the coefficients of employment and migration background are neither statistically significant nor substantial, it can be concluded that they have minimal impact. 

The OLS regression results are likely to be biased because it does not control for unobserved individual differences that can affect the change in concerns for the environment. For example, respondents' political views and personal traits that make them more sensitive to global threats are not controlled for.

```{r results="asis"}
soep_bal_1011$plh0036 <- as.numeric(soep_bal_1011$plh0036)

ols_1 <- lm(plh0036 ~ fukushima, data=soep_bal_1011)
ols_2 <- lm(plh0036 ~ fukushima + sex + agec + emp + edu + mig, data=soep_bal_1011)

stargazer(ols_1, ols_2, type="latex", title = "Ordinary Least Squares", header=FALSE)
```

# Fixed Effects 

In fixed effects model, the effect of unobserved individual heterogeneity cancels out, giving more unbiased results. This model yields a similar result, but the magnitude of the impact is smaller than the OLS: Fukushima changed increased environmental concerns by 0.036 (1.2%). This model is still limited because there may be other external events that canceled out the effect of Fukushima, such as the announcement of nuclear phase-out policy (Goebel et al., 2013). 

Since there is not much time-variance in variables that measure life satisfaction and health, their effects cannot be properly assessed. 

```{r results="asis"}
fe_1 <- plm(plh0036 ~ fukushima, data=soep_bal_1011, index = c("id"), model="within")
fe_2 <- plm(plh0036 ~ fukushima + satis + health, data=soep_bal_1011, index = "id", model = "within")

stargazer(fe_1, fe_2, type="latex", title ="Fixed Effects", header=FALSE)
```


# Difference in Difference

In 


## Data prep for Diff in Diff


```{r results='asis'}
soep_did <- soep_bal_1011 %>% 
  group_by(id) %>% 
  mutate(fukushima = as.numeric(fukushima)) %>%
  mutate(plh0036_lag = dplyr::lag(plh0036),
         fukushima_lag = dplyr::lag(fukushima)) %>%
  ungroup() %>%
  mutate(diff_plh0036 = plh0036 - plh0036_lag,
         diff_fukushima = fukushima - fukushima_lag)

did <- lm(diff_plh0036 ~ diff_fukushima, data=soep_did)
stargazer(did, type="latex", title ="Differenc in Difference", header=FALSE)
```


# References 

Berger, E. M. (2008). A Note on the High Stability of Happiness: The Minimal Effects of a Nuclear Catastrophe on Life Satisfaction. SSRN Electronic Journal. https://doi.org/10.2139/ssrn.1143830 

Goebel, Krekel, Tiefenbach, Ziebarth. (2013). "Natural Disaster, Policy Action, and Mental Well-Being: The Case of Fukushima," IZA Discussion Papers 7691, Institute of Labor Economics (IZA).

Kingdon, J. W. (2011). Agendas, alternatives, and public policies. Boston: Longman.
