---
title: "Explore research questions"
author: "Uni Lee"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(dplyr)
library(ggplot2)
```

# Research question: How does the expansion of renewable energy affect carbon emissions? 

In recent years, renewable energy was rapidly deployed to curb carbon emissions. Has the renewable energy actually decreased carbon emissions? 

Theoretically, it should. According to the "Kaya identity (Kaya and Keiichi, 1997)", emissions are also affected by the number of population, GDP, energy consumption per GDP and carbon intensity of energy. Increased share of renewable energy decreases the carbon intensity of energy. Therefore, it can be hypothesized that the higher the share of renewable energy is, the lower the carbon emissions when the population, GDP and energy consumption per capita are held constant.  

$$
Carbon Emissions = Population * GDP * \frac{Energy}{GDP} * \frac{CO^2}{Energy}
$$

## Data 

To test this assumption, we will use two sources of data:

* Our World in Data, [CO2 and Greenhouse Gas Emissions dataset](https://github.com/owid/co2-data)

This is a country-level panel data on carbon emissions, population, GDP from 1751 to 2018. 

* Our World in Data, [Data on energy](https://github.com/owid/energy-data)).   

This dataset provides national macroeconomic meterics, energy mix, electricity mix and other relevant measures from 1980 to 2019 for 242 countries.

```{r}
energy_raw <- read.csv("data/04_co2/owid-energy-data.csv", sep=";")
names(energy_raw)

co2_raw <- read.csv("data/04_co2/owid-co2-data.csv", sep=",")
names(co2_raw)

energy <- energy_raw %>% select(
  iso_code, country, year,
  gdp, population,
  energy_per_gdp,
  coal_share_elec,
  coal_share_energy,
  gas_share_elec,
  gas_share_energy,
  hydro_share_elec, hydro_share_energy,
  nuclear_share_elec,
  nuclear_share_energy,
  oil_share_elec,
  oil_share_energy,
  renewables_share_elec,
  renewables_share_energy
) %>% na.omit() %>% filter(country !="World")
nrow(energy)

co2 <- co2_raw %>% select(
  country, year,
  co2, co2_growth_prct, co2_per_unit_energy) %>% na.omit()
nrow(co2)

co2_energy <- left_join(energy, co2, by=c("country", "year"))

ggplot(co2_energy, aes(x=renewables_share_energy, y=co2_growth_prct)) +
  geom_point()

ggplot(co2_energy, aes(x=year, y=co2_growth_prct)) +
  geom_point()

```

# Preliminary descriptive analysis 

```{r}

```


# Methodology: fixed effects 




# References 



Kaya, Yoichi; Yokoburi, Keiichi (1997). Environment, energy, and economy : strategies for sustainability. Tokyo [u.a.]: United Nations Univ. Press. ISBN 9280809113.

# Research methods
1) Kaplan-meier: median survival time (how long does it take for something to go through something?) 
2) Log-rank test: are the group differences significant? 
3) Cox: What are the rates of an event occuring and how are they different among groups?
4) DiD / Fixed effects: what is the effect of x on y, controlling for time-constant unobserved heterogeneity?  

- Change in carbon emissions after carbon pricing was introduced 
 => diff-in-diff
  * Controls for time-constant characteristics of the country
  * Cannot control for time-varying heterogeneity 
  * Need a balanced dataset 
  * Control for GDP 


What is the effect of more renewable energy share in carbon emissions?
 => Fixed effects 
 * controls for time-constant characteristics of the country
 * control for time-varying heterogeneity: GDP, population, 
 * longitudinal data for % of renewables in the energy mix and carbon emissions